---
title: "BSA_data_analysis"
author: "Karim-Mane, Fatoumata Seck"
format: docx
---

```{r setup}
#| message: false
#| warning: false


# install {pak}
if (!require("pak")) install.packages("pak")

# install {cleanepi}
if (!require("cleanepi")) pak::pak("epiverse-trace/cleanepi")

# install {rstatix}
if (!require("rstatix")) pak::pak("rstatix")

# install {ggResidpanel}
if (!require("ggResidpanel")) pak::pak("ggResidpanel")

# install {ggpubr}
if (!require("ggpubr")) pak::pak("ggpubr")

# install {patchwork}
if (!require("patchwork")) pak::pak("patchwork")


# load the required libraries
library(cleanepi)
library(readxl)
library(dplyr)
library(magrittr)
library(kableExtra)
library(ggplot2)
library(rstatix)
library(ggResidpanel)
library(ggpubr)
library(patchwork)

# set the path to the dataset
path_to_data <- file.path(dirname(getwd()), "data", "base_fs_km.xlsx")

# load the functions in helpers.R
source(file.path(getwd(), "helpers.R"))
```

# The experiment

The experiment consists in testing whether any of the four chosen inhibitors has
an impact on the _Anopheles reproduction pattern_.

Every inhibitor was tested across three experiments or replicates denoted as
`a`, `b`, and `c`, where one set of mosquitoes was exposed to the inhibitor,
while the other was used as control.

## Data cleaning

```{r}
data <- readxl::read_xlsx(path_to_data)
```

There were `r nrow(data)` females used in the study.

Missing values are written as `na`. We will substitute them with `NA` for
compliance with R.

```{r}
data <- cleanepi::replace_missing_values(
  data,
  target_columns = NULL,
  na_strings = "na"
)
```

After importing the data from MS excel, some `Date` columns have been
transformed into `numeric`. They will be converted back to `Date` in ISO8604
format.
{cleanepi} no longer performs automatic conversion from numeric to Date. We will
use the below to perform the conversion.

```{r}
target_columns <- c("date_ponte", "date_first_eclosion",
                    "date_first_nymphose", "date_last_nymphose",
                    "date_first_emergence", "date_last_emergence")
for (col in target_columns) {
  # convert into numeric
  data[[col]] <- suppressWarnings(
    as.numeric(data[[col]])
  )
  # convert into Date
  data[[col]] <- lubridate::as_date(
    data[[col]],
    origin = as.Date("1900-01-01")
  )
}
```

Rename some column names to make them explicit: replace `experience1` by `inhibitor`, and `experience2` by `experience` as these two columns represent
the inhibitors and experiments that were performed on the corresponding samples.

```{r}
data <- data %>%
  cleanepi::standardize_column_names(
    rename = c(inhibitor = "experience1", experience = "experience2")
  )
```

Rename the values in the two updated columns from the above.

```{r}
# update values in the `inhibitor` column
data <- data %>%
  dplyr::mutate(inhibitor = dplyr::recode(
    inhibitor,
    exp_inhib1 = 'inhib1',
    exp_inhib3 = 'inhib3',
    exp_inhib4 = 'inhib4',
    exp_inhib10 = 'inhib10'
  ))

# update values in the `experience` column
data <- data %>%
  dplyr::mutate(experience = dplyr::recode(
    experience,
    exp1a = 'exp1',
    exp3a = 'exp1',
    exp4a = 'exp1',
    exp10a = 'exp1',
    exp1b = 'exp2',
    exp3b = 'exp2',
    exp4b = 'exp2',
    exp10b = 'exp2',
    exp1c = 'exp3',
    exp3c = 'exp3',
    exp4c = 'exp3',
    exp10c = 'exp3'
  ))

# update values in the `statut` column
are_test <- data$statut != "ctrl"
data$statut[are_test] <- "test"
```

The following columns are not needed for the downstream analysis. We will drop
them.

```{r}
data <- data %>%
  dplyr::select(-c(id, pot, statut2, dose))
```

Moving the `insemination` column to the left to have it as part of the first
columns of the dataset. This makes it accessible when viewing the data in the
View panel.
We also replace 'yes' and 'no' by 'TRUE' and 'FALE' to have the column in a
logical form.

```{r}
data <- data %>%
  dplyr::relocate(insemination, .after = statut) %>%
  dplyr::mutate(insemination = dplyr::recode(
    insemination,
    yes = TRUE,
    no = FALSE
  ))
```

The **insemination** column determines whether a mosquito was inseminated or
not. It contains some missing values where the corresponding mosquito was
actually inseminated. As it's directly related to the eclosion, we will replace every missing entry by 'TRUE' where the **date_first_eclosion** is not missing

```{r}
# identify fertilized females (with hatched eggs) where the insemination value
# is set to NA
incorrect <- is.na(data$insemination) & !is.na(data$date_first_eclosion)
```

we have `r sum(incorrect)` entry at line `r which(incorrect)` where insemination
was set to `NA`, while the females have hatched eggs.

```{r}
# Let's set them to TRUE
if (any(incorrect)) {
  data$insemination[incorrect] <- TRUE
}
```

Some entries has a value in the **date_first_eclosion** column, but their insemination status is set FALSE. We will set it to TRUE to correct for such mistakes.

```{r}
# identify fertilized females (with hatched eggs) where the insemination value
# is set to FALSE
incorrect <- data$insemination == FALSE & !is.na(data$date_first_eclosion)
```

We have `r sum(incorrect)` entries at lines `r which(incorrect)` where the insemination is set to FALSE, while the corresponding females have hatched eggs. 

```{r}
# Let's set them to TRUE
if (any(incorrect)) {
  data$insemination[incorrect] <- TRUE
}
```

## Save the cleaned dataset

After the cleaning steps above, the data is ready for use in analysis workflow.
We are saving this version of the cleaned data that we can share with
collaborators.

```{r}
writexl::write_xlsx(
  data,
  file.path(dirname(path_to_data), "clean_base_fs_km.xlsx")
)
```

## Data analysis

### after investigating the proportion of inseminated females, Remove  non-inseminated females from the target data.

Please specific clearly why you are only interested in inseminated females for downstream analysis.

#Answer (FATOUMATA !!!)
The default conditions for to study reproduction in the lab are :
Male + female Mating, Female's Blood feeding (blood == source of protein )
!!! a proxy a successful mating is the insemination of the female.
During mating male transfer to femelle spermatozoides and other males hormorn. That's what we call  Insemination. 
1. Note that All females selected for this studies are priorly fully blood fled (Condition 2 check !!!)
2. I did reproductive organe dissections after each experiment to check if the females are inseminated. So to ensure Condition 1, we need to select the inseminated females.


```{r}
# convert the insemination column into logical
data$insemination <- as.logical(as.factor(data$insemination))
```

### Comparing the proportion of inseminated females across the different experiences

::: {.callout-note collapse="true"}
## Definition

Inseminated female here refers to the females that have received spermatozoa (mated).
:::

# ! Fatoumata : we should remove data with insemination == na before processing the prop of femelles inseminated because if we select all femelles nbre_femelles = dplyr::n(), and considere inseminees ad insemination == true, so by defautl the na will be consider as insemination == false !!!


```{r message=FALSE}
# generate the table with the proportion of inseminated female in the groups
# defined above (by inhibitor, experience, and statut) Here we remove Insemination == na, as not determined

data <- data %>%
  filter(!is.na(insemination)) # to exclude data where Insemination has not been set


table_insemination <- data %>%
  dplyr::group_by(inhibitor, experience, statut) %>%
  dplyr::summarise(
    nbre_femelles = dplyr::n(),
    inseminees = sum(insemination, na.rm = TRUE) # same sum(insemination == TRUE, na.rm = TRUE)
  ) %>% 
  dplyr::mutate(
    prop_inseminees = (inseminees / nbre_femelles) * 100,
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
        ic_low = purrr::map2_dbl(inseminees, nbre_femelles, ~ prop.test(.x, .y)$conf.int[1] * 100),
        ic_high = purrr::map2_dbl(inseminees, nbre_femelles, ~ prop.test(.x, .y)$conf.int[2] * 100)
    )

# convert to factor and re-level to maintain the order we need
table_insemination$inhibitor <- factor(
  table_insemination$inhibitor,
  levels = c("inhib1", "inhib3", "inhib4", "inhib10")
)

```

Below we are graphically visualising the table with the proportion of inseminated females per group of inhibitor, experience, and statut.

```{r}
plot_prop_inseminated <- ggplot(
  table_insemination,
  aes(
    x = experience,
    y = prop_inseminees,
    fill = statut
    )
) +
  geom_bar(position = 'dodge', stat = 'identity') +
  # Apply renaming and reordering only inside the plot for inhibitors
  facet_wrap(
    ~ factor(
      inhibitor,
      levels = c("inhib1", "inhib3", "inhib4", "inhib10"),
      labels = c("Inhibiteur 1", "Inhibiteur 3", "Inhibiteur 4", "Inhibiteur 10")
    ),
    ncol = 2
  ) +
  # Rename the x-axis tick labels
  scale_x_discrete(
    labels = c(
      "exp1" = "Expérience 1",
      "exp2" = "Expérience 2",
      "exp3" = "Expérience 3"
    )
  ) +
  # Color palette and legend customization
  scale_fill_manual(
    name = "Statut",
    values = c("#0072B2", "#E69F00"),
    labels = c("Contrôle", "Inhibiteur")
  ) +
  labs(
    x = "Expériences",
    y = "Taux d'insémination (%)"
  ) +
  theme_bw() +  # keep default ggplot grid and border
  theme(
    text = element_text(size = 14),
    legend.position = "top",                     # Legend on top
    strip.background = element_rect(fill = "gray80"),  # Gray facet header
    strip.text = element_text(face = "plain")    # Normal facet titles
  ) +
  scale_y_continuous(breaks = seq(0, 100, 15))
# Print the plot
plot_prop_inseminated
```

# Test for the difference in the proportion of inseminated females between the diffent groups

```{r}
prop_test_res <- NULL
for (inhib in unique(table_insemination$inhibitor)) {
  for (exp in unique(table_insemination$experience)) {
    # store the inhibitor and experience information in a data frame
    res <- data.frame(inhibitor = inhib, experience = exp)
    
    # subset the table to only have the data for the experience and inhibitor of
    # interest
    tmp_data <- table_insemination %>%
      dplyr::filter(inhibitor == inhib & experience == exp)
    
    # perform the test
    test_result <- compare_proportions(
      test_data = tmp_data,
      group_col = "statut",
      count_col = "nbre_femelles",
      observed_col = "inseminees"
    )
    
    # select the appropriated columns based the test that was performed
    if (is.null(attr(test_result, "test"))) {
      test_result <- test_result[, c(1, 7, 9:13)]
    } else {
      test_result <- test_result[, c(1, 3:8)]
    }
    res <- cbind(res, test_result)
    
    # combine the data
    prop_test_res <- rbind(prop_test_res, res)
  }
}

```
•	TODO: o	Fatima to analyse the test result in the prop_test_res output.


# Fatoumata : Plot make up, statistical test and labels of Pvalue significance

# choix de test automatique entre Prop_Test, chi2 et fisher
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# 🔹 Mapping numérique pour positionner les barres de p-value
x_mapping <- data.frame(
    experience = c("exp1", "exp2", "exp3"),
    x_num = 1:3
)

# 🔹 Fonction pour choisir automatiquement le test et calculer p-value
# check the pval_data loading afterwards to know the test used for each p
get_pvalue_label_prop_auto <- function(df){
  
  inseminees_ctrl <- sum(df$inseminees[df$statut == "ctrl"])
  total_ctrl     <- sum(df$nbre_femelles[df$statut == "ctrl"])
  inseminees_test <- sum(df$inseminees[df$statut == "test"])
  total_test     <- sum(df$nbre_femelles[df$statut == "test"])
  
  tab <- matrix(c(
    inseminees_ctrl,
    total_ctrl - inseminees_ctrl,
    inseminees_test,
    total_test - inseminees_test
  ), nrow = 2, byrow = TRUE)
  
  colnames(tab) <- c("inséminées", "non_inséminées")
  rownames(tab) <- c("Contrôle", "Inhibiteur")
  
  if(min(tab) < 5){
    p <- fisher.test(tab)$p.value
    test_used <- "Fisher"
  } else if(total_ctrl >= 20 & total_test >= 20){
    p <- prop.test(x = c(inseminees_ctrl, inseminees_test),
                   n = c(total_ctrl, total_test))$p.value
    test_used <- "Prop test"
  } else {
    p <- chisq.test(tab, correct = FALSE)$p.value
    test_used <- "Chi²"
    # En pratique, pour des effectifs ≥ 20, la condition “prop.test” couvre déjà ce que le Chi² ferait, donc on n’utilise plus le Chi² classique dans ce cas précis.
  }
  
  if(p >= 0.05){
    label <- "ns"
  } else if(p < 0.001){
    label <- "***"
  } else if(p < 0.01){
    label <- "**"
  } else {
    label <- "*"
  }
  
  return(c(p = p, label = label, test_used = test_used))
}

# 🔹 Calcul des IC95% (Clopper-Pearson)
ci_data <- table_insemination %>%
    group_by(inhibitor, experience, statut) %>%
    summarise(
        prop = sum(inseminees) / sum(nbre_femelles),
        n = sum(nbre_femelles),
        inseminees = sum(inseminees),
        ci_low = binom.test(sum(inseminees), sum(nbre_femelles))$conf.int[1] * 100,
        ci_high = binom.test(sum(inseminees), sum(nbre_femelles))$conf.int[2] * 100,
        prop_inseminees = prop * 100,
        .groups = "drop"
    )

# 🔹 Calcul des p-values et labels automatiquement
pval_data_prop_insemination <- table_insemination %>%
    group_by(inhibitor, experience) %>%
    summarise(
        tmp = list(get_pvalue_label_prop_auto(cur_data())),
        ci_max = max(
            binom.test(sum(inseminees[statut=="ctrl"]), sum(nbre_femelles[statut=="ctrl"]))$conf.int[2]*100,
            binom.test(sum(inseminees[statut=="test"]), sum(nbre_femelles[statut=="test"]))$conf.int[2]*100
        ),
        .groups = "drop"
    ) %>%
    tidyr::unnest_wider(tmp) %>%
    left_join(x_mapping, by = "experience") %>%
    mutate(
        y.position = ci_max * 1.05,   # barre horizontale au-dessus IC95%
        label_y    = ci_max * 1.12    # label p-value encore plus haut
    )

### 🔹 Affichage du tableau des p-values avec le test utilisé !!! 
pval_data_prop_insemination %>%
    select(inhibitor, experience, p, label, test_used) %>%
    print(n = Inf)

# 🔹 Création du plot
plot_prop_inseminated <- ggplot(
    ci_data,
    aes(x = experience, y = prop_inseminees, fill = statut)
) +
    geom_bar(
        position = position_dodge(width = 0.7), 
        stat = "identity",
        color = "black",
        width = 0.6
    ) +
    geom_errorbar(
        aes(ymin = prop_inseminees, ymax = ci_high),
        position = position_dodge(0.7),
        width = 0.2, 
        linewidth = 0.5,
        color = "black"
    ) +
    facet_wrap(
        ~ factor(
            inhibitor,
            levels = c("inhib1", "inhib10", "inhib3", "inhib4"), #inhib10 will be inhib 2
            labels = c("Inhibiteur 1", "Inhibiteur 2", "Inhibiteur 3", "Inhibiteur 4") 
        ),
        ncol = 2
    ) +
    scale_x_discrete(
        labels = c(
            "exp1" = "Expérience 1",
            "exp2" = "Expérience 2",
            "exp3" = "Expérience 3"
        )
    ) +
    scale_fill_manual(
        name = "Statut",
        values = c("ctrl" = "#0072B2", "test" = "#E69F00"),
        labels = c("Contrôle", "Inhibiteur")
    ) +
    labs(
        x = "Expériences",
        y = "Taux d'insémination (%)"
    ) +
    theme_bw() +
    theme(
        text = element_text(size = 14),
        legend.position = "top",
        strip.background = element_rect(fill = "gray80"),
        axis.title.x = element_text(margin = margin(t = 5)), #margin (t) is to add space on top of the title, betwen the x title and the x axis,
        axis.title.y = element_text(margin = margin(r = 5)), #margin (r) is to add space at the right of the title, betwen the y title and the y axis,
        strip.text = element_text(face = "plain")
    ) +
    scale_y_continuous(breaks = seq(0, 100, 20)) +
  # Barres horizontales p-values bien décalées
    geom_segment(
        data = pval_data_prop_insemination,
        aes(x = x_num - 0.25, xend = x_num + 0.25, y = y.position, yend = y.position),
        inherit.aes = FALSE
    ) +
    geom_text(
        data = pval_data_prop_insemination,
        aes(x = x_num, y = label_y, label = label),
        inherit.aes = FALSE,
        size = 4
    )

# 🔹 Affichage
plot_prop_inseminated

```
```{r}
#Save with more height (inches) 
#ggsave("plot_prop_Insemination.png", plot = plot_prop_inseminated, width = 8, height = 7, dpi = 300)
```

#Lets flip the plot into horizontal bars to see (FS)
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    inseminees_ctrl <- sum(df$inseminees[df$statut == "ctrl"])
    total_ctrl     <- sum(df$nbre_femelles[df$statut == "ctrl"])
    inseminees_test <- sum(df$inseminees[df$statut == "test"])
    total_test     <- sum(df$nbre_femelles[df$statut == "test"])
    
    tab <- matrix(c(
        inseminees_ctrl,
        total_ctrl - inseminees_ctrl,
        inseminees_test,
        total_test - inseminees_test
    ), nrow = 2, byrow = TRUE)
    
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(inseminees_ctrl, inseminees_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    return(c(p = p, label = label, test_used = test_used))
}

# 🔹 Calcul IC95%
ci_data <- table_insemination %>%
    group_by(inhibitor, experience, statut) %>%
    summarise(
        prop = sum(inseminees) / sum(nbre_femelles),
        n = sum(nbre_femelles),
        ci_low = binom.test(sum(inseminees), sum(nbre_femelles))$conf.int[1] * 100,
        ci_high = binom.test(sum(inseminees), sum(nbre_femelles))$conf.int[2] * 100,
        prop_inseminees = prop * 100,
        .groups = "drop"
    )

# 🔹 Calcul p-values
pval_data_prop_insemination <- table_insemination %>%
    group_by(inhibitor, experience) %>%
    summarise(
        tmp = list(get_pvalue_label_prop_auto(cur_data())),
        ci_max = max(
            binom.test(sum(inseminees[statut=="ctrl"]), sum(nbre_femelles[statut=="ctrl"]))$conf.int[2]*100,
            binom.test(sum(inseminees[statut=="test"]), sum(nbre_femelles[statut=="test"]))$conf.int[2]*100
        ),
        .groups = "drop"
    ) %>%
    tidyr::unnest_wider(tmp) %>%
    mutate(
        y_num = as.numeric(factor(experience, levels = c("exp1", "exp2", "exp3"))),
        x.position = ci_max * 1.05,   # position du segment
        label_x    = ci_max * 1.12    # position du label
    )

# 🔹 Graphique avec barres horizontales
#plot_prop_inseminated <- ggplot(
    #ci_data,
    #aes(x = experience, y = prop_inseminees, fill = statut)
plot_prop_inseminated2 <- ggplot(ci_data, aes(
    y = experience, x = prop_inseminees, fill = statut # axis flipped !!!
)) +
    geom_col( # !!!
        position = position_dodge(width = 0.7),
        color = "black", width = 0.6
    ) +
    geom_errorbarh(  # note le "h" pour horizontal
        aes(xmin = ci_low, xmax = ci_high),
        position = position_dodge(0.7),
        width = 0.2,
        linewidth = 0.5,
        color = "black"
    ) +
    facet_wrap(
        ~ factor(
            inhibitor, 
            levels = c("inhib1", "inhib10", "inhib3", "inhib4"),
            labels = c("Inhibiteur 1", "Inhibiteur 2", "Inhibiteur 3", "Inhibiteur 4") #Rename inhib #10 as inhib 2
        ),
        ncol = 2
    ) +
    scale_y_discrete(
        labels = c(
            "exp1" = "Expérience 1",
            "exp2" = "Expérience 2",
            "exp3" = "Expérience 3"
        )
    ) +
    scale_fill_manual(
        name = "Statut",
        values = c("ctrl" = "#0072B2", "test" = "#E69F00"),
        labels = c("Contrôle", "Inhibiteur")
    ) +
    labs(
        x = "Taux d'insémination (%)",
        #y = "Expériences" to remove the title in this case as it a repeat
        y = NULL 
    ) +
    theme_bw() +
    theme(
        text = element_text(size = 14),
        legend.position = "top",
        strip.background = element_rect(fill = "gray80"),
    axis.text.x = element_text(size = 13, color = "black"),
    axis.text.y = element_text(size = 13, color = "black"),   
        axis.title.x = element_text(margin = margin(t = 5)), #margin (t) is to add space on top of the title, betwen the x title and the x axis,
        #axis.title.y = element_text(margin = margin(r = 5)), #margin (r) is to add space at the right of the title, betwen the y title and the y axis,
        strip.text = element_text(face = "plain")
    ) +
    scale_x_continuous(breaks = seq(0, 100, 20)) +
    geom_segment(
        data = pval_data_prop_insemination,
        aes(y = y_num - 0.25, yend = y_num + 0.25, x = x.position, xend = x.position),
        inherit.aes = FALSE
    ) +
    geom_text(
        data = pval_data_prop_insemination,
        aes(y = y_num, x = label_x, label = label),
        inherit.aes = FALSE,
        size = 4
    )

# 🔹 Affichage
plot_prop_inseminated2
```
```{r}
#Save with more height (inches) 
#ggsave("plot_prop_Insemination_ok.png", plot = plot_prop_inseminated2, width = 8, height = 7, dpi = 300)
```

#  Test statistiques sur la proportion de femelles inseminées !!!

```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    inseminees_ctrl <- sum(df$inseminees[df$statut == "ctrl"])
    total_ctrl  <- sum(df$nbre_femelles[df$statut == "ctrl"])
    inseminees_test <- sum(df$inseminees[df$statut == "test"])
    total_test  <- sum(df$nbre_femelles[df$statut == "test"])
    
    prop_ctrl <- (inseminees_ctrl / total_ctrl) * 100
    prop_test <- (inseminees_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(inseminees_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(inseminees_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(inseminees_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(inseminees_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        inseminees_ctrl,
        total_ctrl - inseminees_ctrl,
        inseminees_test,
        total_test - inseminees_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(inseminees_ctrl, inseminees_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Test", "Test > Ctrl"))
    
    tibble(
        Prop = "Inseminees",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_Taux_Insemination <- table_insemination %>%
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
```


#  Global (By Inhib  Ctrl Vs Test) Test statistiques sur la proportion de femelles inseminées !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    inseminees_ctrl <- sum(df$inseminees[df$statut == "ctrl"])
    total_ctrl  <- sum(df$nbre_femelles[df$statut == "ctrl"])
    inseminees_test <- sum(df$inseminees[df$statut == "test"])
    total_test  <- sum(df$nbre_femelles[df$statut == "test"])
    
    prop_ctrl <- (inseminees_ctrl / total_ctrl) * 100
    prop_test <- (inseminees_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(inseminees_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(inseminees_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(inseminees_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(inseminees_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        inseminees_ctrl,
        total_ctrl - inseminees_ctrl,
        inseminees_test,
        total_test - inseminees_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(inseminees_ctrl, inseminees_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Test", "Test > Ctrl"))
    
    tibble(
        Prop = "Inseminees",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_global_Taux_Insemination <- table_insemination %>%
    #group_by(inhibitor, experience) %>%
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
```


-----------------
#Proportion de femelles ayant pondu des oeufs # Here we select insemination == TRUE as explained below

```{r message=FALSE}
# generate the table with the proportion of egg laid females in the groups
# defined above (by inhibitor, experience, and statut)
table_Ayant_pondu <- data %>%
    dplyr::group_by(inhibitor, experience, statut) %>%
    dplyr::summarise(
        #nbre_femelles = dplyr::n(),
        nbre_femelles = sum(insemination, na.rm = TRUE),
        ayant_pondu = sum(insemination == TRUE & nbre_oeufs > 0),
        non_pondu = nbre_femelles - ayant_pondu,
    ) %>% 
    dplyr::mutate(
      prop_ayant_pondu = (ayant_pondu / nbre_femelles * 100),
      prop_non_pondu = (non_pondu / nbre_femelles * 100),
      # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
        ic_low = purrr::map2_dbl(ayant_pondu, nbre_femelles, ~ prop.test(.x, .y)$conf.int[1] * 100),
        ic_high = purrr::map2_dbl(ayant_pondu, nbre_femelles, ~ prop.test(.x, .y)$conf.int[2] * 100)
    )

# convert to factor and re-level to maintain the order we need
table_Ayant_pondu$inhibitor <- factor(
    table_Ayant_pondu$inhibitor,
    levels = c("inhib1", "inhib3", "inhib4", "inhib10")
    )
```


# lets plot the proportion of inseminated females
# Plot including the pourcentages of pondu / non pondu 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

# ---- 1️⃣ Préparation des données ----
table_Ayant_pondu_long <- table_Ayant_pondu %>%
  select(inhibitor, experience, statut, prop_ayant_pondu, prop_non_pondu) %>%
  pivot_longer(
    cols = c(prop_ayant_pondu, prop_non_pondu),
    names_to = "etat_ponte",
    values_to = "proportion"
  ) %>%
  # ✅ Normalisation pour que la somme des proportions = 1 par groupe
  group_by(inhibitor, experience, statut) %>%
  mutate(proportion = proportion / sum(proportion)) %>%
  ungroup() %>%
  mutate(
    etat_ponte = recode(
      etat_ponte,
      "prop_ayant_pondu" = "Ayant pondu",
      "prop_non_pondu" = "N’ayant pas pondu"
    ),
    experience = factor(
      experience,
      levels = c("exp1", "exp2", "exp3"),
      labels = c("Expérience 1", "Expérience 2", "Expérience 3")
    ),
    inhibitor = factor(
      inhibitor,
      levels = c("inhib1", "inhib10", "inhib3", "inhib4"),
      labels = c("Inhibiteur 1", "Inhibiteur 2", "Inhibiteur 3", "Inhibiteur 4") #inhib10 = "Inhibiteur 2" !!!
    ),
    statut = recode(statut,
                    "ctrl" = "Contrôle",
                    "test" = "Inhibiteur"),
    fill_group = paste0(statut, "_", etat_ponte)
  )

# ---- 2️⃣ Couleurs ----
couleurs <- c(
  "Contrôle_Ayant pondu" = "#0072B2",
  "Contrôle_N’ayant pas pondu" = "#99BFE3",
  "Inhibiteur_Ayant pondu" = "#E69F00",
  "Inhibiteur_N’ayant pas pondu" = "#F3C979"
)

# ---- 3️⃣ Graphique ----
plot_Ayant_pondu <- ggplot(table_Ayant_pondu_long, aes(x = statut, y = proportion, fill = fill_group)) +
  geom_col(width = 0.8, color = "white", position = position_stack()) +
  # ✅ Pourcentages arrondis à l’unité #remove this geom_text () to not display the %
  geom_text(
            aes(label = percent(proportion, accuracy = 1)),
            position = position_stack(vjust = 0.5),
            size = 2.8,
            color = "black"
            ) +
  coord_polar(theta = "y") +
  facet_grid(experience ~ inhibitor) +
  scale_fill_manual(
    name = "Proportion de femelles",
    values = couleurs,
    labels = c(
      "Contrôle_Ayant pondu" = "Contrôle – Ayant pondu",
      "Contrôle_N’ayant pas pondu" = "Contrôle – N’ayant pas pondu",
      "Inhibiteur_Ayant pondu" = "Inhibiteur – Ayant pondu",
      "Inhibiteur_N’ayant pas pondu" = "Inhibiteur – N’ayant pas pondu"
    ),
    guide = guide_legend(
      nrow = 2,
      byrow = TRUE,
      override.aes = list(color = "black")  # ✅ encadrer les carrés de légende
    )
  ) +
  labs(x = NULL, y = NULL) +
  theme_bw(base_size = 13) +
  theme(
    legend.position = "top",
    legend.title = element_text(face = "bold",color = "black", size = 12),
    legend.text = element_text(size = 10, color = "black"),
    strip.background = element_rect(fill = "gray95", color = "gray80"),
    strip.text = element_text(face = "plain", color = "black"),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    panel.spacing.x = unit(1.2, "lines"),
    panel.spacing.y = unit(1, "lines")
  )

# ---- 4️⃣ Affichage ----
plot_Ayant_pondu
```

```{r}
#Save with more height (inches) 
#ggsave("plot_prop_ayant_pondu_ok.png", plot = plot_Ayant_pondu, width = 8, height = 6, dpi = 300)
```

#  Test statistiques sur la proportion de femelles ayant pondu !!!

```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    ayant_Pondu_ctrl <- sum(df$ayant_pondu[df$statut == "ctrl"])
    total_ctrl  <- sum(df$nbre_femelles[df$statut == "ctrl"])
    ayant_Pondu_test <- sum(df$ayant_pondu[df$statut == "test"])
    total_test  <- sum(df$nbre_femelles[df$statut == "test"])
    
    prop_ctrl <- (ayant_Pondu_ctrl / total_ctrl) * 100
    prop_test <- (ayant_Pondu_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(ayant_Pondu_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(ayant_Pondu_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(ayant_Pondu_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(ayant_Pondu_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        ayant_Pondu_ctrl,
        total_ctrl - ayant_Pondu_ctrl,
        ayant_Pondu_test,
        total_test - ayant_Pondu_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(ayant_Pondu_ctrl, ayant_Pondu_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Test", "Test > Ctrl"))
    
    tibble(
        Prop = "Ayant_Pond",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_Taux_ayant_Pondu <- table_Ayant_pondu %>%
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
```


#  Global (By Inhibitor  Ctrl Vs Test) Test statistiques sur la proportion de femelles ayant pondu !!!

```{r message=FALSE}
# generate the table with the proportion of egg laid females in the groups
# defined above (by inhibitor, experience, and statut)
table_global_Ayant_pondu <- data %>%
    dplyr::group_by(inhibitor, statut) %>%
    dplyr::summarise(
        #nbre_femelles = dplyr::n(),
        nbre_femelles = sum(insemination, na.rm = TRUE),
        ayant_pondu = sum(insemination == TRUE & nbre_oeufs > 0),
        non_pondu = nbre_femelles - ayant_pondu,
    ) %>% 
    dplyr::mutate(
      prop_ayant_pondu = (ayant_pondu / nbre_femelles * 100),
      prop_non_pondu = (non_pondu / nbre_femelles * 100),
      # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
        ic_low = purrr::map2_dbl(ayant_pondu, nbre_femelles, ~ prop.test(.x, .y)$conf.int[1] * 100),
        ic_high = purrr::map2_dbl(ayant_pondu, nbre_femelles, ~ prop.test(.x, .y)$conf.int[2] * 100)
    )

# convert to factor and re-level to maintain the order we need
table_global_Ayant_pondu$inhibitor <- factor(
    table_global_Ayant_pondu$inhibitor,
    levels = c("inhib1", "inhib10", "inhib3", "inhib4")
    )
```
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    ayant_Pondu_ctrl <- sum(df$ayant_pondu[df$statut == "ctrl"])
    total_ctrl  <- sum(df$nbre_femelles[df$statut == "ctrl"])
    ayant_Pondu_test <- sum(df$ayant_pondu[df$statut == "test"])
    total_test  <- sum(df$nbre_femelles[df$statut == "test"])
    
    prop_ctrl <- (ayant_Pondu_ctrl / total_ctrl) * 100
    prop_test <- (ayant_Pondu_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(ayant_Pondu_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(ayant_Pondu_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(ayant_Pondu_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(ayant_Pondu_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        ayant_Pondu_ctrl,
        total_ctrl - ayant_Pondu_ctrl,
        ayant_Pondu_test,
        total_test - ayant_Pondu_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(ayant_Pondu_ctrl, ayant_Pondu_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Test", "Test > Ctrl"))
    
    tibble(
        Prop = "Ayant_Pond",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_global_Taux_ayant_Pondu <- table_Ayant_pondu %>%
    #group_by(inhibitor, experience) %>%
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
```

##### Inhibitors effect on mosquitoe fecundity and cycle length !!!!

The section below focus on understanding the impact of the inhibitors on:

*   the mosquitoes fecundity (number of laid eggs) and
*   the length of their reproduction cycle

The fecundity is represented by the variable `nbre_oeufs`.
The length of mosquitoes reproduction cycle is represented by the
`duree_p_emergence` variable.

Given that only inseminated mosquitoes will have a reproduction cycle that can
generate offspring, we filtered the data such that only inseminated mosquitoes
will be considered in the downstream analyses. We also removed all entries in
the dataset where the number of laid eggs is not greater than 0 as We are just interested in females that have laid eggs in order to fellow their reproduction params 

```{r}
# only keep the inseminated and eggs laid females
data_for_analysis <- data %>%
  dplyr::filter(insemination & nbre_oeufs > 0) #insemination == TRUE
```
```{r}
hist(data_for_analysis$nbre_oeufs, 100)
ggplot2::ggplot(
  data_for_analysis, aes(nbre_oeufs, fill = statut)) + 
  geom_histogram() +
  # facet_wrap( ~ experience, ncol = 2) +
  facet_grid(experience ~ inhibitor)
  theme_bw()
```

The distribution looks quite normal. However the data are on different scales,
with the values ranging between `r range(data_for_analysis$nbre_oeufs)[1]` and
`r range(data_for_analysis$nbre_oeufs)[2]`. To remove the effect of different
scales on the downstream analysis, we will normalise the data using the
**z-score** method.


```{r}
# calculate z-score by inhibitor separately
data_for_analysis$normalised_by_inhib <- ave(
  data_for_analysis$nbre_oeufs,
  data_for_analysis$inhibitor,
  FUN = scale
)

# calculate log values
data_for_analysis$log_normalised <- ave(
  data_for_analysis$nbre_oeufs,
  data_for_analysis$inhibitor,
  FUN = function(x) log(x)
)
```
# histogramme de la distribution du nbre d'oeufs normalisée
```{r}
library(dplyr)
library(ggplot2)

# --- Recodage des variables ---
data_plot <- data_for_analysis %>%
    mutate(
        # Renommer statut
        statut = recode(statut,
                        "ctrl" = "Contrôle",
                        "test" = "Inhibiteur"),
        
        # Recode inhibitor : inhib10 -> inhib2
        inhibitor = recode(inhibitor,
                           "inhib10" = "inhib2"),
        
        # Renommer inhibitor en Inhibiteur 1-4
        inhibitor = recode(inhibitor,
                           "inhib1" = "Inhibiteur 1",
                           "inhib2" = "Inhibiteur 2",
                           "inhib3" = "Inhibiteur 3",
                           "inhib4" = "Inhibiteur 4"),
        
        # Renommer experience
        experience = recode(experience,
                            "exp1" = "Expérience 1",
                            "exp2" = "Expérience 2",
                            "exp3" = "Expérience 3")
    )

# --- Création de l'histogramme ---
histogram_oeufs <- ggplot(data_plot, aes(x = nbre_oeufs, fill = statut)) +
    geom_histogram(color = "black", bins = 30, alpha = 0.8) +
    facet_grid(experience ~ inhibitor) +
    scale_fill_manual(
        name = "Statut",
        values = c("Contrôle" = "#0072B2", "Inhibiteur" = "#E69F00")
    ) +
    scale_x_continuous(
        breaks = seq(0, 300, 50),
        name = "Nombre d'œufs"
    ) +
    scale_y_continuous(
        breaks = seq(0, 8, 2)
    ) +
    labs(
        y = "Fréquence"
    ) +
    theme_bw(base_size = 16) +
    theme(
        text = element_text(family = "serif"),
        axis.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 12),
        axis.text.x = element_text(angle = 35, hjust = 1),   # pour éviter chevauchement
        strip.text = element_text(size = 14, face = "bold"),
        legend.position = "top",
        legend.title = element_text(face = "bold", size = 14),
        legend.text = element_text(size = 14),
        legend.box = "horizontal",
        legend.box.just = "center",
        plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
    )

# --- Afficher le graphique ---
histogram_oeufs
```
```{r}
#Save with more height (inches) 
#ggsave("histogram_nbre_oeufs_ok.png", plot = histogram_oeufs, width = 8, height = 7, dpi = 300)
```

#Mean number of eggs 

```{r}
Table_mean_eggs <- data_for_analysis %>%
    dplyr::group_by(inhibitor, experience, statut) %>%
    summarise(
        nbre_femelles = dplyr::n(),
        total_oeufs = sum(nbre_oeufs, na.rm = TRUE),
        moyenne = mean(nbre_oeufs, na.rm = TRUE),
        ecart_type = sd(nbre_oeufs, na.rm = TRUE),  # La dispersion des valeurs autours de la moyenne 
        moyenne_sd = ecart_type / sqrt(nbre_femelles),  #erreur standard de la moyenne (moyenne_sd) = écart-type/√n, #donne plus de precision à l'estimation de la moyenne.
    )

```

# Number of eggs laid by females 
There are `r nrow(data_for_analysis)` inseminated females. The distribution of
the number of laid eggs per inhibitor is shown in the figure below.

The figure below show the distribution of the number of laid eggs over the delay
of egg laying for each inhibitor.

```{r}
# transform the data of interest in long format
data_for_analysis$duree_g_ponte <- as.numeric(data_for_analysis$duree_g_ponte)

ggplot(data_for_analysis,
       aes(x = duree_g_ponte, y = nbre_oeufs, color = statut)) +
  geom_point() +
  facet_grid(experience ~ inhibitor) +
  theme_bw()
```

```{r}
plot_ponte <- ggplot2::ggplot(
  data_for_analysis, aes(x = experience, y = nbre_oeufs, fill = statut)) + 
  geom_boxplot() +
  stat_summary(
    fun = mean, geom = "point", shape = 20, size = 3, color = "white",
    position = position_dodge2(width = 0.75, preserve = "single")
  ) +
  facet_wrap( ~ inhibitor, ncol = 2) +
  theme_bw()

plot_ponte
```


# Plot make up + pvalues mean differences and significance by FS !!!

```{r} 
library(dplyr)
library(ggplot2)

# Mapping numérique pour l'axe x
x_mapping <- data.frame(
    experience = c("exp1", "exp2", "exp3"),
    x_num = 1:3
)

# Fonction pour calculer p et le label de signification
get_pvalue_label <- function(df){
    ctrl_vals <- df$nbre_oeufs[df$statut=="ctrl"]
    test_vals <- df$nbre_oeufs[df$statut=="test"]
    
    # Vérification normalité
    normal_ctrl <- shapiro.test(ctrl_vals)$p.value > 0.05
    normal_test <- shapiro.test(test_vals)$p.value > 0.05
    
    # Test statistique approprié
    if(normal_ctrl & normal_test){
        p <- t.test(nbre_oeufs ~ statut, data=df)$p.value
        test_used <- "t.test"
    } else {
        #p <- wilcox.test(nbre_oeufs ~ statut, data=df)$p.value
        p <- wilcox.test(nbre_oeufs ~ statut, data=df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    }
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
    return(c(p = p, label = label, test_used = test_used))
}

# Calcul des p-values et labels pour chaque inhibiteur x expérience
pval_data_mean_egg <- data_for_analysis %>%
    group_by(inhibitor, experience) %>%
    summarise(
        tmp = list(get_pvalue_label(cur_data())),
        max_y = max(nbre_oeufs, na.rm = TRUE),
        .groups = "drop"
    ) %>%
    tidyr::unnest_wider(tmp) %>%     # séparer p et label
    left_join(x_mapping, by="experience") %>%
    mutate(
        y.position = max_y * 1.05,          # barre horizontale pour p value
        label_y = y.position + max_y*0.095   # label légèrement au-dessus
    )

# Plot avec barre horizontale et labels significatifs
plot_ponte <- ggplot(data_for_analysis,
                     aes(x = experience, y = nbre_oeufs, fill = statut)) +
    geom_boxplot(color = "black", position = position_dodge(width = 0.8)) +
    facet_wrap(
        ~ factor(
            inhibitor,
            levels = c("inhib1", "inhib10", "inhib3", "inhib4"),
            labels = c("Inhibiteur 1", "Inhibiteur 2", "Inhibiteur 3", "Inhibiteur 4") #inhib10 = "Inhibiteur 2" !!!
        ),
        ncol = 2
    ) +
    scale_x_discrete(
        labels = c(
            "exp1" = "Expérience 1",
            "exp2" = "Expérience 2",
            "exp3" = "Expérience 3"
        )
    ) +
    scale_fill_manual(
        name = "Statut",
        values = c("ctrl" = "#0072B2", "test" = "#E69F00"),
        labels = c("Contrôle", "Inhibiteur")
    ) +
    labs(x = "Expériences", y = "Nombre d'œufs") +
    theme_bw() +
    theme(
        text = element_text(size = 14),
        legend.position = "top",
        strip.background = element_rect(fill = "gray80"),
        axis.title.x = element_text(margin = margin(t = 5)), #margin (t) is to add space on top of the title, betwen the x title and the x axis,
        axis.title.y = element_text(margin = margin(r = 5)), #margin (r) is to add space at the right of the title, betwen the y title and the y axis,
        strip.text = element_text(face = "plain")
    ) +
    # Barre horizontale Ctrl/Test
    geom_segment(
        data = pval_data_mean_egg,
        aes(x = x_num - 0.25, xend = x_num + 0.25, y = y.position, yend = y.position),
        inherit.aes = FALSE
    ) +
    # Label significatif (ns, *, **, ***) légèrement au-dessus
    geom_text(
        data = pval_data_mean_egg,
        aes(x = x_num, y = label_y, label = label),
        inherit.aes = FALSE,
        size = 4
    ) +
scale_y_continuous(breaks = seq(0, 200, 100))

plot_ponte

```
# plot with Mean number of eggs by FS !

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# --- Paramètres de positionnement (ajustables) ---
dodge_width <- 0.8   # doit être le même pour boxplot et pour les points
box_width <- 0.6     # largeur des boxplots visuelles
offset <- 0.22       # décalage numérique pour tracer les segments p-value (≈ dodge_width/2)

# --- Assurer les niveaux d'experience (ordre) ---
#data_for_analysis <- data_for_analysis %>%
    #mutate(experience = factor(experience, levels = c("exp1","exp2","exp3")))

# --- Mean per group (correct placement) ---
mean_data <- data_for_analysis %>%
    group_by(inhibitor, experience, statut) %>%
    summarise(mean_eggs = mean(nbre_oeufs, na.rm = TRUE), .groups = "drop")

# --- Prepare pval_data with numeric x positions (for correct segment placement) ---
# we assume you have pval_data_mean_egg as before (with x_num numeric 1..3)
pval_data2 <- pval_data_mean_egg %>%
    # numeric center position already in x_num; compute left/right for ctrl/test segments
    mutate(
        x_ctrl = x_num - offset,
        x_test = x_num + offset
    )

# --- THE PLOT ---
plot_ponte <- ggplot(data_for_analysis,
                     aes(x = experience, y = nbre_oeufs, fill = statut)) +
    
    # boxplots (dodge so ctrl & test side-by-side)
    geom_boxplot(color = "black",
                 position = position_dodge(width = dodge_width),
                 width = box_width) +
    
    # mean point for each box: use mean_data and same dodge
    geom_point(
        data = mean_data,
        aes(x = experience, y = mean_eggs, fill = statut),
        position = position_dodge(width = dodge_width),
        shape = 16, size = 3.5, stroke = 0.6,
        color = "black"
    ) +
    
    facet_wrap(
        ~ factor(
            inhibitor,
            levels = c("inhib1", "inhib10", "inhib3", "inhib4"),
            labels = c("Inhibiteur 1", "Inhibiteur 2", "Inhibiteur 3", "Inhibiteur 4")
        ),
        ncol = 2
    ) +
    
    scale_x_discrete(
        labels = c("exp1" = "Expérience 1",
                   "exp2" = "Expérience 2",
                   "exp3" = "Expérience 3")
    ) +
    
    scale_fill_manual(
        name = "Statut",
        values = c("ctrl" = "#0072B2", "test" = "#E69F00"),
        labels = c("Contrôle", "Inhibiteur")
    ) +
    
    labs(x = "Expériences", y = "Nombre d'œufs") +
    theme_bw() +
    theme(
        text = element_text(size = 14),
        legend.position = "top",
        strip.background = element_rect(fill = "gray80"),
        axis.title.x = element_text(margin = margin(t = 5)),
        axis.title.y = element_text(margin = margin(r = 5)),
        strip.text = element_text(face = "plain")
    ) +
    
    # P-value bar: draw between the two boxes using the numeric x positions computed
    geom_segment(
        data = pval_data2,
        aes(x = x_ctrl, xend = x_test, y = y.position, yend = y.position),
        inherit.aes = FALSE
    ) +
    
  # vertical small ticks at ends (optional) to look like a bracket)
    #geom_segment(
       # data = pval_data2,
       # aes(x = x_ctrl, xend = x_ctrl, y = y.position, yend = y.position - (max_y * 0.02)),
        #inherit.aes = FALSE
    #) +
    #geom_segment(
        #data = pval_data2,
       # aes(x = x_test, xend = x_test, y = y.position, yend = y.position - (max_y * 0.02)),
       # inherit.aes = FALSE
    #) +
    
# Label for significance (centered)
    geom_text(
        data = pval_data2,
        aes(x = x_num, y = label_y, label = label),
        inherit.aes = FALSE,
        size = 4
    ) +
    
    scale_y_continuous(breaks = seq(0, 200, 100))

# --- Afficher ---
plot_ponte
```


```{r}
#Save with more height (inches) 
#ggsave("plot_nbre_oeufs_ok.png", plot = plot_ponte, width = 8, height = 6.5, dpi = 300)
```

#Global Mean number of eggs 

```{r}
Table_global_mean_eggs <- data_for_analysis %>%
    dplyr::group_by(inhibitor, statut) %>%
    summarise(
        nbre_femelles = dplyr::n(),
        total_oeufs = sum(nbre_oeufs, na.rm = TRUE),
        moyenne = mean(nbre_oeufs, na.rm = TRUE),
        ecart_type = sd(nbre_oeufs, na.rm = TRUE),  # La dispersion des valeurs autours de la moyenne 
        moyenne_sd = ecart_type / sqrt(nbre_femelles), #ESM = sd/# erreur standard de la moyenne (mean precision)
        .groups = "drop"
    )

```

#Global Mean number of eggs Pvals (group by inhibitors)
```{r} 
library(dplyr)
library(tidyr)
library(tibble)

get_pvalue_label <- function(df){

    # Extract values
    ctrl_vals <- df$nbre_oeufs[df$statut == "ctrl"]
    test_vals <- df$nbre_oeufs[df$statut == "test"]

    # Check normality
    normal_ctrl <- shapiro.test(ctrl_vals)$p.value > 0.05
    normal_test <- shapiro.test(test_vals)$p.value > 0.05
    
    # Choose the appropriate test
    if(normal_ctrl & normal_test){
        p <- t.test(nbre_oeufs ~ statut, data = df)$p.value
        test_used <- "t.test"
    } else {
        p <- wilcox.test(nbre_oeufs ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    }

    # Choose significance label
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }

    # Direction of effect
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)

    if(mean_test > mean_ctrl){
        direction <- "test > ctrl"
    } else if(mean_test < mean_ctrl){
        direction <- "ctrl > test"
    } else {
        direction <- "equal"
    }

    return(tibble(
        stade = "oeufs",
        mean_ctrl = mean_ctrl,
        mean_test = mean_test,
        p = p,
        label = label,
        test_used = test_used,
        direction = direction
    ))
}



pval_global_mean_eggs <- data_for_analysis %>%
    group_by(inhibitor) %>%
    #group_by(inhibitor, experience) %>%
    summarise(
        res = list(get_pvalue_label(cur_data())),
        .groups = "drop"
    ) %>%
    unnest(res)
```

#-----------------------------------------------------------

```{r}
# select the data for inhibitor 1
inhibitor1 <- data_for_analysis %>%
  dplyr::filter(inhibitor == "inhib1")

# test for the normality of the number of laid eggs
lm_num_eggs <- lm(nbre_oeufs ~ 1, data = inhibitor1)

# create the diagnostic plot, indicating a QQ-plot
lm_num_eggs %>% 
  ggResidpanel::resid_panel(plots = "qq")

# Shapiro-Wilk test
inhibitor1 %>% 
  rstatix::shapiro_test(nbre_oeufs)

# test the difference in mean number of laid eggs between control and test mosquitoes in every experience
stat_test <- inhibitor1 %>%
  dplyr::group_by(experience) %>%
  rstatix::t_test(nbre_oeufs ~ statut) #%>%
  # rstatix::adjust_pvalue(method = "bonferroni") %>%
  # rstatix::add_significance("p.adj")


# Create a box plot
bxp_inhib1 <- ggpubr::ggboxplot(
  inhibitor1, x = "experience", y = "nbre_oeufs", 
  color = "statut", palette = c("#00AFBB", "#E7B800"),
  add = "mean"
)

# Add p-values onto the box plots
stat_test <- stat_test %>%
  add_xy_position(x = "experience", dodge = 0.8)
bxp_inhib1 + ggpubr::stat_pvalue_manual(
  stat_test,  label = "p", tip.length = 0
)

# Add 10% spaces between the p-value labels and the plot border
bxp_inhib1 + ggpubr::stat_pvalue_manual(
  stat_test,  label = "p", tip.length = 0
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
```


```{r}
# select the data for inhibitor 4
inhibitor4 <- data_for_analysis %>%
  dplyr::filter(inhibitor == "inhib4")

# test for the normality of the number of laid eggs
lm_num_eggs <- lm(nbre_oeufs ~ 1, data = inhibitor4)

# create the diagnostic plot, indicating a QQ-plot
lm_num_eggs %>%
  ggResidpanel::resid_panel(plots = "qq")

# Shapiro-Wilk test
inhibitor4 %>% 
  rstatix::shapiro_test(nbre_oeufs)

# test the difference in mean number of laid eggs between control and test mosquitoes in every experience
stat_test <- inhibitor4 %>%
  dplyr::group_by(experience) %>%
  rstatix::wilcox_test(nbre_oeufs ~ statut) #%>%
  # rstatix::adjust_pvalue(method = "bonferroni") %>%
  # rstatix::add_significance("p.adj")


# Create a box plot
bxp_inhib4 <- ggpubr::ggboxplot(
  inhibitor4, x = "experience", y = "nbre_oeufs", 
  color = "statut", palette = c("#00AFBB", "#E7B800"),
  add = "mean"
)

# Add p-values onto the box plots
stat_test <- stat_test %>%
  add_xy_position(x = "experience", dodge = 0.8)
bxp_inhib4 + ggpubr::stat_pvalue_manual(
  stat_test,  label = "p", tip.length = 0
)

# Add 10% spaces between the p-value labels and the plot border
bxp_inhib4 + ggpubr::stat_pvalue_manual(
  stat_test,  label = "p", tip.length = 0
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))
```

```{r}
# put the plots together
bxp_inhib1 + bxp_inhib4
```
-----------------------------------------------------------
----------------------------------------------------------

## Inhibitors effect on mosquitoe fertility and eclosion delay

variable with the number of larves: `nbre_larves`
variable with the eclosion delay: `duree_p_eclosion`
variable with the fertility rate: `taux_eclosion`

```{r}
data_for_analysis$taux_eclosion <- as.numeric(data_for_analysis$taux_eclosion)
data_for_analysis$taux_eclosion <- round(data_for_analysis$taux_eclosion,
                                         digits = 2)

# filter out rows where the duration of eclosion is NA due to missing values in
# either the 'date_ponte' or 'date_eclosion' column
data_for_analysis %>%
  #dplyr::filter(!is.na(duree_p_eclosion)) %>%
  ggplot(aes(x = duree_p_eclosion, y = taux_eclosion, color = statut)) +
  # geom_violin(trim = FALSE) +  # Use trim = TRUE to cut off the tails
  # geom_boxplot(width = 0.1, position = position_dodge(0.9)) #, color = "black"
  geom_jitter(width = 0.2, height = 0.2, size = 1) +
  geom_point() +
  facet_grid(experience ~ inhibitor) +
  theme_bw()
```

---------------

## Inhibitors effect on mosquitoe nymphose rate

variable with the nymphose length: `duree_p_nymphose`
variable with the nymphose rate: `taux_nymphose`

```{r}
data_for_analysis$taux_nymphose <- as.numeric(data_for_analysis$taux_nymphose)
data_for_analysis$taux_nymphose <- round(data_for_analysis$taux_nymphose,
                                         digits = 2)
data_for_analysis$duree_p_nymphose <- as.numeric(data_for_analysis$duree_p_nymphose)

test <- data_for_analysis %>%
  dplyr::group_by(experience, inhibitor, statut) %>%
  dplyr::count(duree_p_nymphose)

# time distribution only
data_for_analysis %>%
  dplyr::filter(nbre_larves > 0) %>%
  ggplot(aes(x = duree_p_nymphose, color = statut)) +
  geom_histogram(alpha = 0.5, position="identity") +
  facet_grid(experience ~ inhibitor) +
  theme_bw() +
  scale_x_continuous(breaks = sort(unique(data_for_analysis$duree_p_nymphose))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
# distribution of rate over time
data_for_analysis %>%
  dplyr::filter(nbre_larves > 0) %>%
  ggplot(aes(x = duree_p_nymphose, y = taux_nymphose, color = statut)) +
  geom_jitter(width = 0.2, height = 0.2, size = 1) +
  facet_grid(experience ~ inhibitor) +
  theme_bw()
```

## Inhibitors effect on mosquitoe emergence rate

variable with the emergence length: `duree_p_emergence`
variable with the emergence rate: `taux_emergence`

```{r}
# taux emergence is the ratio between the number of adults and the nymphe
# but this need to filter out all mosquitoes with 0 number of nymphe
data_for_analysis$taux_emergence <- round(as.numeric(
  data_for_analysis$nbre_adultes / data_for_analysis$nbre_nymphes
) * 100, digits = 2)
data_for_analysis$duree_p_emergence <- as.numeric(
  data_for_analysis$duree_p_emergence
)

# distribution of rate over time
data_for_analysis %>%
  dplyr::filter(nbre_nymphes > 0) %>%  
ggplot(aes(x = duree_p_emergence, y = taux_emergence, color = statut)) +
  geom_point() +
  # geom_jitter(width = 0.2, height = 0.2, size = 1) +
  facet_grid(experience ~ inhibitor) +
  theme_bw() +
  scale_y_continuous(breaks = seq(0, 100, 10))
```

#Add the mean duration and emergence_rate line for each (Fatoumata !!!)
```{r}
# Compute the mean values per group
mean_lines <- data_for_analysis %>%
  dplyr::filter(nbre_nymphes > 0) %>%  
  group_by(statut, experience, inhibitor) %>%
  summarise(
    mean_x = mean(duree_p_emergence, na.rm = TRUE),
    mean_y = mean(taux_emergence, na.rm = TRUE)
  ) %>%
  ungroup()

# Then plot
ggplot(data_for_analysis, aes(x = duree_p_emergence, y = taux_emergence, color = statut)) +
  
  # smaller data points
  geom_point(size = 1) +

  # vertical mean line (dashed)
  geom_segment(data = mean_lines,
               aes(x = mean_x, xend = mean_x, y = 0, yend = mean_y, color = statut),
               linetype = "dashed", size = 0.5) +

  # horizontal mean line (dashed)
  geom_segment(data = mean_lines,
               aes(x = 0, xend = mean_x, y = mean_y, yend = mean_y, color = statut),
               linetype = "dashed", size = 0.5) +

  # mean point (intersection, larger)
  geom_point(data = mean_lines, aes(x = mean_x, y = mean_y, color = statut),
             shape = 18, size = 4) +  # size increased for visibility

  facet_grid(experience ~ inhibitor) +
  theme_bw() +
  scale_y_continuous(breaks = seq(0, 100, 10)) +
  scale_x_continuous(breaks = seq(0, 20, 5))

```


## Variation of mean numbers FS !

```{r message=FALSE}
phase_ponte <- data_for_analysis %>% 
  dplyr::filter(nbre_oeufs > 0) %>%
  dplyr::group_by(experience, inhibitor, statut) %>% 
  summarise(
    moyenne = mean(nbre_oeufs, na.rm = TRUE),
    ecart_type = sd(nbre_oeufs, na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()), #erreur_standard de la moyenne = écart-type/√n, #precision de la moyenne
    stade = "oeufs"
  )

phase_eclosion <- data_for_analysis %>% 
  dplyr::filter(nbre_larves > 0) %>%  
  dplyr::group_by(experience, inhibitor, statut) %>% 
  summarise(
    moyenne = mean(nbre_larves, na.rm = TRUE),
     ecart_type = sd(nbre_larves, na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()), #erreur_standard, #precision de la moyenne
    stade = "larves"
  )

phase_nymphose <- data_for_analysis %>%
  dplyr::filter(nbre_nymphes > 0) %>%  
  dplyr::group_by(experience, inhibitor, statut) %>% 
  summarise(
    moyenne = mean(nbre_nymphes, na.rm = TRUE),
     ecart_type = sd(nbre_nymphes, na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()), #erreur_standard, #precision de la moyenne
    stade = "nymphes"
  )

phase_emergence <- data_for_analysis %>% 
  dplyr::filter(nbre_adultes > 0) %>%  
  dplyr::group_by(experience, inhibitor, statut) %>% 
  summarise(
    moyenne = mean(nbre_adultes, na.rm = TRUE),
     ecart_type = sd(nbre_adultes, na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()), #erreur_standard, #precision de la moyenne
    stade = "adultes"
  )

resume_cycle <- rbind(
  phase_ponte, phase_eclosion, phase_nymphose, phase_emergence
)
```

#Labels legends (Fatoumata)
```{r}
resume_cycle$inhibitor <- factor(
  resume_cycle$inhibitor,
  levels = c("inhib1", "inhib10", "inhib3", "inhib4")
)

resume_cycle$stade <- factor(
  resume_cycle$stade,
  levels = c("oeufs", "larves", "nymphes", "adultes")
)

plot_mean_dev_cycle <- ggplot(resume_cycle, aes(x = stade, y = moyenne, color = statut, group = statut)) +
  geom_line(size = 0.75) +
  geom_point(size = 2.5) +
  labs(x = "Stades de développement", y = "Nombres moyens", color = "Statut") +
  theme_minimal() +
  facet_grid(
    experience ~ inhibitor,
    labeller = labeller(
      inhibitor = c(
        inhib1 = "Inhibiteur 1",
        inhib10 = "Inhibiteur 2",
        inhib3 = "Inhibiteur 3",
        inhib4 = "Inhibiteur 4"
      ),
      experience = c(
        exp1 = "Expérience 1",
        exp2 = "Expérience 2",
        exp3 = "Expérience 3"
      )
    )
  ) +
  scale_color_manual(
    name = "Statut", 
    values = c("ctrl" = "#0072B2", "test" = "#E69F00"),
    labels = c("ctrl" = "Contrôle", "test" = "Inhibiteur")
  ) +
  theme(
    text = element_text(size = 12),
    axis.text.x = element_text(size = 12, angle = 35, hjust = 1, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),                 
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 7)), #margin (t) is to add space on top of the title, betwen the x title and the x axis
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 7)), #margin (t) is to add space at right of the title, betwen the x title and the x axis,
    legend.position = "top",
    legend.title = element_text(size = 12, face = "bold", color = "black"),
    legend.text = element_text(size = 12, face = "bold"),
    strip.text = element_text(face = "bold", size = 12) # facet titles
  ) +
  scale_y_continuous(breaks = seq(0, 120, 20))
plot_mean_dev_cycle
```
```{r}
#Save with more height (inches) 
#ggsave("plot_mean_dev_cycle_ok.png", plot = plot_mean_dev_cycle, width = 8, height = 6, dpi = 300)
```
##Fatoumata_Curiosity !!!!
## Variation of mean durations
```{r}
# conform the data of interest are into long format
data_for_analysis$duree_g_ponte <- as.numeric(data_for_analysis$duree_g_ponte)
data_for_analysis$duree_p_eclosion <- as.numeric(data_for_analysis$duree_p_eclosion)
data_for_analysis$duree_p_nymphose <- as.numeric(data_for_analysis$duree_p_nymphose)
data_for_analysis$duree_p_emergence <- as.numeric(data_for_analysis$duree_p_emergence)
```
```{r message=FALSE}
delay_ponte <- data_for_analysis %>%  
  dplyr::filter(nbre_oeufs > 0) %>%  
  dplyr::group_by(inhibitor,experience, statut) %>%  
  summarise(
    moyenne = mean(duree_g_ponte, na.rm = TRUE),
    ecart_type = sd(duree_g_ponte, na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    phase = "ponte"
  )

delay_eclosion <- data_for_analysis %>%  
  dplyr::filter(nbre_larves > 0) %>%  
  dplyr::group_by(inhibitor,experience, statut) %>%  
  summarise(
    moyenne = mean((duree_g_ponte+duree_p_eclosion), na.rm = TRUE),
    ecart_type = sd((duree_g_ponte+duree_p_eclosion), na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    phase = "eclosion"
  )

delay_nymphose <- data_for_analysis %>%  
  dplyr::filter(nbre_nymphes > 0) %>%  
  dplyr::group_by(inhibitor,experience, statut) %>%  
  summarise(
    moyenne = mean((duree_g_ponte+duree_p_nymphose), na.rm = TRUE),
    ecart_type = sd((duree_g_ponte+duree_p_nymphose), na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    phase = "nymphose"
  )

delay_emergence <- data_for_analysis %>%  
  dplyr::filter(nbre_adultes > 0) %>%  
  dplyr::group_by(inhibitor,experience, statut) %>%  
  summarise(
    moyenne = mean((duree_g_ponte+duree_p_emergence), na.rm = TRUE),
    ecart_type = sd((duree_g_ponte+duree_p_emergence), na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    phase = "emergence"
  )

resume_cycle_durations <- rbind(
  delay_ponte, delay_eclosion, delay_nymphose, delay_emergence
)
```

```{r}
resume_cycle_durations$phase <- factor(
  resume_cycle_durations$phase,
  levels = c("ponte", "eclosion", "nymphose", "emergence")
)

resume_cycle_durations$inhibitor <- factor(
  resume_cycle_durations$inhibitor,
  levels = c("inhib1","inhib10", "inhib3", "inhib4")
)

plot_duration_dev_cycle <- ggplot(resume_cycle_durations, aes(x = phase, y = moyenne, color = statut, group = statut)) +
 geom_line(size = 0.75) +
  geom_point(size = 2.5) +
  labs(x = "Etapes du cycle de développement", y = "Nombres moyens de jours", color = "Statut") +
  theme_minimal() +
  facet_grid(
    experience ~ inhibitor,
    labeller = labeller(
      inhibitor = c(
        inhib1 = "Inhibiteur 1",
        inhib3 = "Inhibiteur 3",
        inhib10 = "Inhibiteur 2",
        inhib4 = "Inhibiteur 4"
      ),
      experience = c(
        exp1 = "Expérience 1",
        exp2 = "Expérience 2",
        exp3 = "Expérience 3"
      )
    )
  ) +
  scale_color_manual(
    values = c("ctrl" = "#0072B2", "test" = "#E69F00"),
    labels = c("ctrl" = "Contrôle", "test" = "Inhibiteur")
  ) +
  theme(
    text = element_text(size = 13),
    axis.text.x = element_text(size = 13, angle = 35, hjust = 1, color = "black"),
    axis.text.y = element_text(size = 13, color = "black", margin = margin(t = 12)), #margin (t) is to add space on top of the y axis elements
    axis.title.x = element_text(size = 13, face = "bold", margin = margin(t = 7)), #margin (t) is to add space on top of the title, betwen the x title and the x axis
    axis.title.y = element_text(size = 13, face = "bold", margin = margin(r = 7)), #margin (t) is to add space at right of the title, betwen the x title and the x axis,
    legend.position = "top",
    legend.title = element_text(size = 13, face = "bold", color = "black"),
    legend.text = element_text(size = 13, face = "bold"),
    strip.text = element_text(face = "bold", size = 13) # facet titles
  ) +
  scale_y_continuous(breaks = seq(0, 20, 2), expand = expansion(mult = c(0.05, 0.1)))
plot_duration_dev_cycle
```
```{r}
#Save with more height (inches) 
#ggsave("plot_duration_dev_cycle_ok.pdf", plot = plot_duration_dev_cycle, width = 8, height = 7, dpi = 300)
```
```{r}
ggplot(resume_cycle, aes(x = stade, y = moyenne, group = statut)) +
  # geom_line(alpha = 0.6) +
  # geom_point(size = 1.8, alpha = 0.8) +
  geom_smooth(aes(color = statut, fill = statut),
              method = "lm", se = TRUE, linewidth = 0.9, alpha = 0.20) +  # per-sample lm
  labs(x = "Cycle", y = "nombre moyen", color = "Statut",
       title = "Per-sample trends with linear fits") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
resume_cycle
```
# effect of inhibitors on the sex-ratio
```{r message=FALSE}
#Prepartion de la table
table_sexe <- data_for_analysis %>%
    dplyr::filter(nbre_adultes > 0) %>%  
    dplyr::group_by(inhibitor, experience, statut) %>%
    dplyr::summarise(
        adultes = sum(nbre_adultes, na.rm = TRUE),
        males = sum(nbre_males, na.rm = TRUE),
        femelles = sum(nbre_femelles, na.rm = TRUE),) %>% 
    dplyr::mutate(
        prop_femelles = (femelles / adultes) * 100,
        prop_males = (males / adultes) * 100
    )

# convert to factor and re-level to maintain the order we need
table_sexe$inhibitor <- factor(
    table_sexe$inhibitor,
    levels = c("inhib1", "inhib3", "inhib4", "inhib10")
)
```

```{r}
#creation du plot
library(dplyr)
library(tidyr)
library(ggplot2)

# Convert to long format ----
table_sexe_long <- table_sexe %>%
    tidyr::pivot_longer(
        cols = c(prop_femelles, prop_males),
        names_to = "Sexe",
        values_to = "proportion"
    ) %>%
    dplyr::mutate(
        Sexe = dplyr::recode(Sexe,
                             "prop_femelles" = "Femelles",
                             "prop_males" = "Mâles"),
        Sexe = factor(Sexe, levels = c("Mâles", "Femelles")), # Femelles bottom
        experience = factor(experience,
                            levels = c("exp1", "exp2", "exp3"),
                            labels = c("Expérience 1", "Expérience 2", "Expérience 3")),
        statut = dplyr::recode(statut,
                               "ctrl" = "Contrôle",
                               "test" = "Inhibiteur"),
        inhibitor = factor(inhibitor,
                           levels = c("inhib1", "inhib10", "inhib3", "inhib4"),
                           labels = c("Inhibiteur 1", "Inhibiteur 2", "Inhibiteur 3", "Inhibiteur 4"))
    )

# ---- 3️⃣ Refined, thesis-ready color palette ----
couleurs <- c(
    "Femelles" = "#E7B800", #"sienna3", #"tomato3",
    "Mâles" = "#00AFBB"  #"olivedrab"
)

# ---- 4️⃣ Plot ----
plot_sex_ratio <- ggplot(table_sexe_long, aes(
    x = statut,
    y = proportion,
    fill = Sexe
)) +
    geom_col(position = "stack", width = 0.7, color = "black") +
    geom_hline(yintercept = 50, color = "red", linewidth = 0.60) +  # solid elegant green line
    facet_grid(experience ~ inhibitor) +
    scale_fill_manual(
        values = couleurs,
        guide = guide_legend(
            override.aes = list(color = "black")  # Outline legend boxes
        )
    ) +
    scale_y_continuous(
        name = "Proportion (%)",
        limits = c(0, 100),
        breaks = seq(0, 100, 25),
        labels = function(x) paste0(x, "%")
    ) +
    labs(x = NULL) +
    theme_bw(base_size = 13) +
    theme(
        legend.position = "top",
        legend.title = element_text(face = "bold", size = 11, color = "black"),
        legend.text = element_text(size = 11, color = "black"),
        strip.background = element_rect(fill = "gray95", color = "gray80", linewidth = 0.8),
        strip.text = element_text(face = "plain", size = 12, color = "black"),
        axis.text = element_text(size = 11, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        panel.grid.major = element_line(color = "gray90"),
        panel.grid.minor = element_blank(),
        panel.spacing.x = unit(1.2, "lines"),
        panel.spacing.y = unit(1, "lines"),
        text = element_text(color = "black")
    )

# ---- 5️⃣ Display ----
plot_sex_ratio

```
```{r}
#Save with more height (inches) 
#ggsave("plot_Sex_ratio_ok.png", plot = plot_sex_ratio, width = 8, height = 7, dpi = 300)
```

#  Test statistiques sur la proportion de femelles emergentes !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    femelles_ctrl <- sum(df$femelles[df$statut == "ctrl"])
    total_ctrl     <- sum(df$adultes[df$statut == "ctrl"])
    femelles_test <- sum(df$femelles[df$statut == "test"])
    total_test     <- sum(df$adultes[df$statut == "test"])
    
    prop_ctrl <- (femelles_ctrl / total_ctrl) * 100
    prop_test <- (femelles_test / total_test) * 100
    
     # calcul des IC à 95 % avec prop.test()
        
         ic_low_ctrl = purrr::map2_dbl(femelles_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(femelles_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(femelles_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(femelles_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        femelles_ctrl,
        total_ctrl - femelles_ctrl,
        femelles_test,
        total_test - femelles_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(femelles_ctrl, femelles_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(prop_ctrl > prop_test, "Ctrl > Test", "Test > Ctrl"))
    
    tibble(
        Prop = "Femelles",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_Emergence_femelles <- table_sexe %>%
    group_by(inhibitor, experience) %>%
    #group_by(inhibitor) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
pval_Emergence_femelles
```

#  Global Test statistiques sur la proportion de femelles emergentes !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    femelles_ctrl <- sum(df$femelles[df$statut == "ctrl"])
    total_ctrl     <- sum(df$adultes[df$statut == "ctrl"])
    femelles_test <- sum(df$femelles[df$statut == "test"])
    total_test     <- sum(df$adultes[df$statut == "test"])
    
    prop_ctrl <- (femelles_ctrl / total_ctrl) * 100
    prop_test <- (femelles_test / total_test) * 100
    
     # calcul des IC à 95 % avec prop.test()
        
         ic_low_ctrl = purrr::map2_dbl(femelles_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(femelles_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(femelles_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(femelles_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        femelles_ctrl,
        total_ctrl - femelles_ctrl,
        femelles_test,
        total_test - femelles_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(femelles_ctrl, femelles_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(prop_ctrl > prop_test, "Ctrl > Test", "Test > Ctrl"))
    
    tibble(
        Prop = "Femelles",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

Global_pval_Emergence_femelles <- table_sexe %>%
    #group_by(inhibitor, experience) %>%
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
Global_pval_Emergence_femelles
```

---------------------------------------------------------
#Globale Summary_Inhibitors_Analyse (Fatoumata)


#1. Lets get the means number of each stade of devoloppement, summarized by inhibitors and status
 
```{r message=FALSE}
Summary_oeufs <- data_for_analysis %>% 
    dplyr::filter(nbre_oeufs > 0) %>% 
    dplyr::group_by(inhibitor, statut) %>% 
    summarise(
       nbre_femelles = dplyr::n(),
        moyenne = mean(nbre_oeufs, na.rm = TRUE),
    ecart_type = sd(nbre_oeufs, na.rm = TRUE),  #variabilité entre individus
    moyenne_sd = ecart_type / sqrt(dplyr::n()), #erreur_standard, #precision de la moyenne
        stade = "oeufs"
    )

Summary_Larves <- data_for_analysis %>% 
    dplyr::filter(nbre_larves > 0) %>% 
    dplyr::group_by(inhibitor, statut) %>% 
    summarise(
       nbre_femelles = dplyr::n(),
        moyenne = mean(nbre_larves, na.rm = TRUE),
    ecart_type = sd (nbre_larves, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()), 
        stade = "larves"
    )

Summary_Nymphes <- data_for_analysis %>% 
    dplyr::filter(nbre_nymphes > 0) %>% 
    dplyr::group_by(inhibitor, statut) %>% 
    summarise(
       nbre_femelles = dplyr::n(),
        moyenne = mean(nbre_nymphes, na.rm = TRUE),
     ecart_type = sd (nbre_nymphes, na.rm = TRUE), 
     moyenne_sd = ecart_type / sqrt(dplyr::n()),
        stade = "nymphes"
    )

Summary_Adultes <- data_for_analysis %>% 
    dplyr::filter(nbre_adultes > 0) %>% 
    dplyr::group_by(inhibitor, statut) %>% 
    summarise(
        nbre_femelles = dplyr::n(),
        moyenne = mean(nbre_adultes, na.rm = TRUE),
  ecart_type = sd (nbre_adultes, na.rm = TRUE), 
  moyenne_sd = ecart_type / sqrt(dplyr::n()),
        stade = "adultes"
    )

Summary_global_means_cycle <- rbind(
    Summary_oeufs, Summary_Larves, Summary_Nymphes, Summary_Adultes
)
```


#  Taux d'eclosion and Test statistiques sur les taux d'eclosion par exp !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    eclos_ctrl <- sum(df$nbre_larves[df$statut == "ctrl"], na.rm = TRUE)
    total_ctrl  <- sum(df$nbre_oeufs[df$statut == "ctrl"], na.rm = TRUE)
    eclos_test <- sum(df$nbre_larves[df$statut == "test"], na.rm = TRUE)
    total_test  <- sum(df$nbre_oeufs[df$statut == "test"], na.rm = TRUE)
    
    prop_ctrl <- (eclos_ctrl / total_ctrl) * 100
    prop_test <- (eclos_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(eclos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(eclos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(eclos_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(eclos_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        eclos_ctrl ,
        total_ctrl - eclos_ctrl ,
        eclos_test,
        total_test - eclos_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(eclos_ctrl, eclos_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Inhib", "Inhib > Ctrl"))
    
    tibble(
        Phase = "Taux_Eclos",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_Taux_eclosion <- data_for_analysis %>%
    dplyr::filter(nbre_oeufs > 0) %>% 
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
pval_Taux_eclosion
```

#  Taux de nymphose and Test statistiques sur les taux de nymphose par exp !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    nymphos_ctrl <- sum(df$nbre_nymphes[df$statut == "ctrl"], na.rm = TRUE)
    total_ctrl  <- sum(df$nbre_larves[df$statut == "ctrl"], na.rm = TRUE)
    nymphos_test <- sum(df$nbre_nymphes[df$statut == "test"], na.rm = TRUE)
    total_test  <- sum(df$nbre_larves[df$statut == "test"], na.rm = TRUE)
    
    prop_ctrl <- (nymphos_ctrl / total_ctrl) * 100
    prop_test <- (nymphos_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(nymphos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(nymphos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(nymphos_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(nymphos_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        nymphos_ctrl ,
        total_ctrl - nymphos_ctrl ,
        nymphos_test,
        total_test - nymphos_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(nymphos_ctrl, nymphos_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Inhib", "Inhib > Ctrl"))
    
    tibble(
        Phase = "Taux_Nymph",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_Taux_Nymphose <- data_for_analysis %>%
    dplyr::filter(nbre_larves > 0) %>% 
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
pval_Taux_Nymphose
```

#  Taux de Emergence and Test statistiques sur les taux de Emergence par exp !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    emerg_ctrl <- sum(df$nbre_adultes[df$statut == "ctrl"], na.rm = TRUE) #emergeant
    total_ctrl  <- sum(df$nbre_nymphes[df$statut == "ctrl"], na.rm = TRUE) #total
    emerg_test<- sum(df$nbre_adultes[df$statut == "test"], na.rm = TRUE)
    total_test  <- sum(df$nbre_nymphes[df$statut == "test"], na.rm = TRUE)
    
    prop_ctrl <- (emerg_ctrl / total_ctrl) * 100
    prop_test <- (emerg_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(emerg_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(emerg_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(emerg_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(emerg_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        emerg_ctrl ,
        total_ctrl - emerg_ctrl ,
        emerg_test,
        total_test - emerg_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(emerg_ctrl, emerg_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Inhib", "Inhib > Ctrl"))
    
    tibble(
        Phase = "Taux_Emerg",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_Taux_Emergence <- data_for_analysis %>%
    dplyr::filter(nbre_nymphes > 0) %>% 
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
pval_Taux_Emergence
```

#  Taux d'eclosion and Test statistiques sur les taux d'eclosion global par inhibitor !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    eclos_ctrl <- sum(df$nbre_larves[df$statut == "ctrl"], na.rm = TRUE)
    total_ctrl  <- sum(df$nbre_oeufs[df$statut == "ctrl"], na.rm = TRUE)
    eclos_test <- sum(df$nbre_larves[df$statut == "test"], na.rm = TRUE)
    total_test  <- sum(df$nbre_oeufs[df$statut == "test"], na.rm = TRUE)
    
    prop_ctrl <- (eclos_ctrl / total_ctrl) * 100
    prop_test <- (eclos_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(eclos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(eclos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(eclos_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(eclos_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        eclos_ctrl ,
        total_ctrl - eclos_ctrl ,
        eclos_test,
        total_test - eclos_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(eclos_ctrl, eclos_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Inhib", "Inhib > Ctrl"))
    
    tibble(
        Phase = "Taux_Eclos",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_global_Taux_eclosion <- data_for_analysis %>%
    dplyr::filter(nbre_oeufs > 0) %>% 
    #group_by(inhibitor, experience) %>%
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
pval_global_Taux_eclosion
```

#  Taux de nymphose and Test statistiques sur les taux de nymphose par exp !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    nymphos_ctrl <- sum(df$nbre_nymphes[df$statut == "ctrl"], na.rm = TRUE)
    total_ctrl  <- sum(df$nbre_larves[df$statut == "ctrl"], na.rm = TRUE)
    nymphos_test <- sum(df$nbre_nymphes[df$statut == "test"], na.rm = TRUE)
    total_test  <- sum(df$nbre_larves[df$statut == "test"], na.rm = TRUE)
    
    prop_ctrl <- (nymphos_ctrl / total_ctrl) * 100
    prop_test <- (nymphos_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(nymphos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(nymphos_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(nymphos_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(nymphos_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        nymphos_ctrl ,
        total_ctrl - nymphos_ctrl ,
        nymphos_test,
        total_test - nymphos_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(nymphos_ctrl, nymphos_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Inhib", "Inhib > Ctrl"))
    
    tibble(
        Phase = "Taux_Nymph",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_global_Taux_Nymphose <- data_for_analysis %>%
    dplyr::filter(nbre_larves > 0) %>% 
    #group_by(inhibitor, experience) %>%
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
pval_global_Taux_Nymphose
```

#  Taux de Emergence and Test statistiques sur les taux de Emergence par exp !!!
```{r}
# 🔹 Fonction automatique de choix de test
get_pvalue_label_prop_auto <- function(df){
    emerg_ctrl <- sum(df$nbre_adultes[df$statut == "ctrl"], na.rm = TRUE) #emergeant
    total_ctrl  <- sum(df$nbre_nymphes[df$statut == "ctrl"], na.rm = TRUE) #total
    emerg_test<- sum(df$nbre_adultes[df$statut == "test"], na.rm = TRUE)
    total_test  <- sum(df$nbre_nymphes[df$statut == "test"], na.rm = TRUE)
    
    prop_ctrl <- (emerg_ctrl / total_ctrl) * 100
    prop_test <- (emerg_test / total_test) * 100
    
    # calcul des IC à 95 % avec prop.test() #Fatoumata !!!
    ic_low_ctrl = purrr::map2_dbl(emerg_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_ctrl = purrr::map2_dbl(emerg_ctrl, total_ctrl, ~ prop.test(.x, .y)$conf.int[2] * 100)
    ic_low_test = purrr::map2_dbl(emerg_test, total_test, ~ prop.test(.x, .y)$conf.int[1] * 100)
    ic_high_test = purrr::map2_dbl(emerg_test, total_test, ~ prop.test(.x, .y)$conf.int[2] * 100)
    
    tab <- matrix(c(
        emerg_ctrl ,
        total_ctrl - emerg_ctrl ,
        emerg_test,
        total_test - emerg_test
    ), nrow = 2, byrow = TRUE)
    
    # Choix du test selon les conditions
    if(min(tab) < 5){
        p <- fisher.test(tab)$p.value
        test_used <- "Fisher"
    } else if(total_ctrl >= 20 & total_test >= 20){
        p <- prop.test(x = c(emerg_ctrl, emerg_test),
                       n = c(total_ctrl, total_test))$p.value
        test_used <- "Prop test"
    } else {
        p <- chisq.test(tab, correct = FALSE)$p.value
        test_used <- "Chi²"
    }
    
    # Label de significativité
    label <- ifelse(p >= 0.05, "ns",
                    ifelse(p < 0.001, "***",
                           ifelse(p < 0.01, "**", "*")))
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                        ifelse(prop_ctrl > prop_test, "Ctrl > Inhib", "Inhib > Ctrl"))
    
    tibble(
        Phase = "Taux_Emerg",
        prop_ctrl = round(prop_ctrl, 3),
        ic_low_ctrl = round(ic_low_ctrl, 3),
        ic_high_ctrl = round(ic_high_ctrl, 3),
        prop_test = round(prop_test, 3),
        ic_low_test = round(ic_low_test, 3),
        ic_high_test = round(ic_high_test, 3),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used,
    )
}

pval_global_Taux_Emergence <- data_for_analysis %>%
    dplyr::filter(nbre_nymphes > 0) %>% 
    #group_by(inhibitor, experience) %>%
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label_prop_auto(pick(everything())),
        .groups = "drop"
    )
pval_global_Taux_Emergence
```

#2. Lets get the means duration between of each stade of devoloppement, summarized by experiences,inhibitors and status
 
```{r message=FALSE}
Sum_duree_G_Ponte <- data_for_analysis %>%
  dplyr::filter(nbre_oeufs > 0) %>% 
  #dplyr::group_by(inhibitor, statut) %>% 
  dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean(duree_g_ponte, na.rm = TRUE),
   ecart_type = sd(duree_g_ponte, na.rm = TRUE), #variabilité entre individus
   moyenne_sd = ecart_type / sqrt(dplyr::n()), #erreur_standard, #precision de la moyenne
   #moyenne_sd = sd(duree_g_ponte, na.rm = TRUE) / sqrt(sum(!is.na(duree_g_ponte))) #if ecart_type not needed
    durée = "Gorgement_Ponte"
  )

Sum_duree_P_Eclosion <- data_for_analysis %>%
  dplyr::filter(nbre_larves > 0) %>% 
  #dplyr::group_by(inhibitor, statut) %>% 
  dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_p_eclosion), na.rm = TRUE),
    ecart_type = sd(duree_p_eclosion, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Ponte_Eclosion"
  )

data_for_analysis2 <- data_for_analysis %>%
    mutate(
        duree_eclo_nymphose = duree_p_nymphose - duree_p_eclosion
    )

Sum_duree_Eclo_Nymphose <- data_for_analysis2 %>% 
  dplyr::filter(nbre_nymphes > 0) %>% 
  #dplyr::group_by(inhibitor, statut) %>% 
  dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_eclo_nymphose), na.rm = TRUE),
    ecart_type = sd(duree_eclo_nymphose, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Eclosion_Nymphose"
  )

data_for_analysis <- data_for_analysis %>%
        mutate(
            duree_n_emergence = as.numeric(duree_n_emergence)
        )

Sum_duree_N_Emergence<- data_for_analysis %>% 
  dplyr::filter(nbre_adultes > 0) %>% 
  #dplyr::group_by(inhibitor, statut) %>% 
  dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_n_emergence), na.rm = TRUE),
    ecart_type = sd(duree_n_emergence, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Nymphose_Emergence"
  )

Sum_duree_P_Emergence<- data_for_analysis %>% 
  dplyr::filter(nbre_adultes > 0) %>% 
  dplyr::group_by(inhibitor, statut) %>% 
  #dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_p_emergence), na.rm = TRUE),
    ecart_type = sd(duree_p_emergence, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Ponte_Emergence"
  )

Summary_cycle_durations <- rbind(
  Sum_duree_G_Ponte, Sum_duree_P_Eclosion, Sum_duree_Eclo_Nymphose, Sum_duree_N_Emergence, Sum_duree_P_Emergence
)
```

#3. Lets do the statiscal test on the means difference delay between stade of devoloppement, summarized by experience, inhibitors and status (Ctrl Vs Test for each inhitor)
# Duration Gorgement_Ponte 
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df)
  {
    ctrl_vals <- df$duree_g_ponte[df$statut == "ctrl"]
    test_vals <- df$duree_g_ponte[df$statut == "test"]
    
     mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
     mean_test <- mean(test_vals, na.rm = TRUE)
    
    ### Because of Shapiro error message (duree can be = 1 so < 2), knowing we the appropried test is the Wilcoxon, #we do not need to loop !!!
    
#--- Test de Wilcoxon directement ---
        p <- wilcox.test(duree_g_ponte ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                     )
    
    tibble(durée = "Gorgement_Ponte",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
       Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_duree_g_ponte <- data_for_analysis %>%
    dplyr::filter(nbre_oeufs > 0) %>% 
    #group_by(inhibitor) %>%
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
# For Ponte_Eclosion
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df) {
    
    ctrl_vals <- df$duree_p_eclosion[df$statut == "ctrl"]
    test_vals <- df$duree_p_eclosion[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
    ### Because of Shapiro error message (duree can be = 1 so < 2), knowing we the appropried test is the Wilcoxon, 
    #we do not need to loop !!!
  #--- Test de Wilcoxon directement ---
        p <- wilcox.test(duree_p_eclosion ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
        
  # --- Significance label ---
  label <- if (is.na(p)) {
    NA
  } else if (p >= 0.05) {
    "ns"
  } else if (p < 0.001) {
    "***"
  } else if (p < 0.01) {
    "**"
  } else {
    "*"
  }
    
    # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                     )
    
    tibble(durée = "Ponte_Eclosion",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
       Signif = Signif,
        test_used = test_used
    )
}
# --- Calcul des p-values pour chaque inhibiteur ---
pval_duree_p_eclosion <- data_for_analysis %>%
    dplyr::filter(nbre_larves > 0) %>% 
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
# For Eclosion_Nymphose 
```{r}
data_for_analysis2 <- data_for_analysis %>%
    mutate(
        duree_eclo_nymphose = duree_p_nymphose - duree_p_eclosion
    )
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_eclo_nymphose[df$statut == "ctrl"]
    test_vals <- df$duree_eclo_nymphose[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
    ### Because of Shapiro error message (duree can be = 1 so < 2), knowing we the appropried test is the Wilcoxon, 
#we do not need to loop !!!
#--- Test de Wilcoxon directement ---  
    p <- wilcox.test(duree_eclo_nymphose ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                     )
    
    tibble(durée = "Eclosion_Nymphose",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_duree_eclo_nymphose <- data_for_analysis2 %>%
    dplyr::filter(nbre_nymphes > 0) %>% 
    #group_by(inhibitor) %>%
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
# For Nymphose_Emergence 
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_n_emergence[df$statut == "ctrl"]
    test_vals <- df$duree_n_emergence[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
  ### Because of Shapiro error message, knowing we the appropried test is the Wilcoxon, 
    #we do not the loop !!!
  #--- Test de Wilcoxon directement ---
        p <- wilcox.test(duree_n_emergence ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                     )
    
    
    tibble(durée = "Nymphose_Emergence",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
       Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_duree_n_emergence <- data_for_analysis %>%
    dplyr::filter(nbre_adultes > 0) %>% 
    #group_by(inhibitor) %>%
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```

# For Ponte_Emergence 
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_p_emergence[df$statut == "ctrl"]
    test_vals <- df$duree_p_emergence[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
      ### Because of Shapiro error message (duree can be = 1 so < 2), knowing we the appropried test is the Wilcoxon, 
#we do not need to loop !!!
#--- Test de Wilcoxon directement ---
        p <- wilcox.test(duree_p_emergence ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
      
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                     )
    
    tibble(durée = "Ponte_Emergence",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_duree_p_emergence <- data_for_analysis %>%
    dplyr::filter(nbre_adultes > 0) %>% 
    #group_by(inhibitor) %>%
    group_by(inhibitor, experience) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
```{r}
Summary_pval_durations_cycle <- rbind(
  pval_duree_g_ponte,
  pval_duree_p_eclosion,
  pval_duree_eclo_nymphose,
  pval_duree_n_emergence,
  pval_duree_p_emergence
)
```



#4. Lets get the means duration between of each stade of devoloppement, summarized by inhibitors and status
 
```{r message=FALSE}
Sum_duree_G_Ponte <- data_for_analysis %>%
  dplyr::filter(nbre_oeufs > 0) %>% 
  dplyr::group_by(inhibitor, statut) %>% 
  #dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean(duree_g_ponte, na.rm = TRUE),
   ecart_type = sd(duree_g_ponte, na.rm = TRUE), #variabilité entre individus
   moyenne_sd = ecart_type / sqrt(dplyr::n()), #erreur_standard, #precision de la moyenne
   #moyenne_sd = sd(duree_g_ponte, na.rm = TRUE) / sqrt(sum(!is.na(duree_g_ponte))) #if ecart_type not needed
    durée = "Gorgement_Ponte"
  )

Sum_duree_P_Eclosion <- data_for_analysis %>%
  dplyr::filter(nbre_larves > 0) %>% 
  dplyr::group_by(inhibitor, statut) %>% 
  #dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_p_eclosion), na.rm = TRUE),
    ecart_type = sd(duree_p_eclosion, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Ponte_Eclosion"
  )

data_for_analysis2 <- data_for_analysis %>%
    mutate(
        duree_eclo_nymphose = duree_p_nymphose - duree_p_eclosion
    )

Sum_duree_Eclo_Nymphose <- data_for_analysis2 %>% 
  dplyr::filter(nbre_nymphes > 0) %>% 
  dplyr::group_by(inhibitor, statut) %>% 
  #dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_eclo_nymphose), na.rm = TRUE),
    ecart_type = sd(duree_eclo_nymphose, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Eclosion_Nymphose"
  )

data_for_analysis <- data_for_analysis %>%
        mutate(
            duree_n_emergence = as.numeric(duree_n_emergence)
        )

Sum_duree_N_Emergence<- data_for_analysis %>% 
  dplyr::filter(nbre_adultes > 0) %>% 
  dplyr::group_by(inhibitor, statut) %>% 
  #dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_n_emergence), na.rm = TRUE),
    ecart_type = sd(duree_n_emergence, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Nymphose_Emergence"
  )

Sum_duree_P_Emergence<- data_for_analysis %>% 
  dplyr::filter(nbre_adultes > 0) %>% 
  dplyr::group_by(inhibitor, statut) %>% 
  #dplyr::group_by(inhibitor, experience, statut) %>% 
  summarise(
    moyenne = mean((duree_p_emergence), na.rm = TRUE),
    ecart_type = sd(duree_p_emergence, na.rm = TRUE), 
    moyenne_sd = ecart_type / sqrt(dplyr::n()),
    durée = "Ponte_Emergence"
  )

Summary_global_cycle_durations <- rbind(
  Sum_duree_G_Ponte, Sum_duree_P_Eclosion, Sum_duree_Eclo_Nymphose, Sum_duree_N_Emergence, Sum_duree_P_Emergence
)
```

#5. Lets do the statiscal test on the means difference delay between stade of devoloppement, summarized by inhibitors and status (Ctrl Vs Test for each inhitor)
# Duration Gorgement_Ponte 
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_g_ponte[df$statut == "ctrl"]
    test_vals <- df$duree_g_ponte[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
      # Vérification normalité
    normal_ctrl <- shapiro.test(ctrl_vals)$p.value > 0.05
    normal_test <- shapiro.test(test_vals)$p.value > 0.05
    
    # Test statistique approprié
    if(normal_ctrl & normal_test){
        p <- t.test(duree_g_ponte ~ statut, data = df)$p.value
        test_used <- "t.test"
    } else {
        p <- wilcox.test(duree_g_ponte ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    }
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                    )
    
    tibble(durée = "Gorgement_Ponte",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_global_duree_g_ponte <- data_for_analysis %>%
    dplyr::filter(nbre_oeufs > 0) %>% 
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
# For Ponte_Eclosion
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_p_eclosion[df$statut == "ctrl"]
    test_vals <- df$duree_p_eclosion[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
      # Vérification normalité
    normal_ctrl <- shapiro.test(ctrl_vals)$p.value > 0.05
    normal_test <- shapiro.test(test_vals)$p.value > 0.05
    
    # Test statistique approprié
    if(normal_ctrl & normal_test){
        p <- t.test(duree_p_eclosion ~ statut, data = df)$p.value
        test_used <- "t.test"
    } else {
        p <- wilcox.test(duree_p_eclosion ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    }
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                    )
    
    tibble(durée = "Ponte_Eclosion",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_global_duree_p_eclosion <- data_for_analysis %>%
    dplyr::filter(nbre_larves > 0) %>% 
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
# For Eclosion_Nymphose 
```{r}
data_for_analysis2 <- data_for_analysis %>%
    mutate(
        duree_eclo_nymphose = duree_p_nymphose - duree_p_eclosion
    )
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_eclo_nymphose[df$statut == "ctrl"]
    test_vals <- df$duree_eclo_nymphose[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
      # Vérification normalité
    normal_ctrl <- shapiro.test(ctrl_vals)$p.value > 0.05
    normal_test <- shapiro.test(test_vals)$p.value > 0.05
    
    # Test statistique approprié
    if(normal_ctrl & normal_test){
        p <- t.test(duree_eclo_nymphose ~ statut, data = df)$p.value
        test_used <- "t.test"
    } else {
        p <- wilcox.test(duree_eclo_nymphose ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    }
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                    )
    
    tibble(durée = "Eclosion_Nymphose",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_global_duree_eclo_nymphose <- data_for_analysis2 %>%
    dplyr::filter(nbre_nymphes > 0) %>% 
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
# For Nymphose_Emergence 
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_n_emergence[df$statut == "ctrl"]
    test_vals <- df$duree_n_emergence[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
      # Vérification normalité
    normal_ctrl <- shapiro.test(ctrl_vals)$p.value > 0.05
    normal_test <- shapiro.test(test_vals)$p.value > 0.05
    
    # Test statistique approprié
    if(normal_ctrl & normal_test){
        p <- t.test(duree_n_emergence ~ statut, data = df)$p.value
        test_used <- "t.test"
    } else {
        p <- wilcox.test(duree_n_emergence ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    }
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                    )
    
    
    tibble(durée = "Nymphose_Emergence",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_global_duree_n_emergence <- data_for_analysis %>%
    dplyr::filter(nbre_adultes > 0) %>% 
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```

# For Ponte_Emergence 
```{r}
# --- Fonction pour calculer p-value, test et label ---
get_pvalue_label <- function(df){
    ctrl_vals <- df$duree_p_emergence[df$statut == "ctrl"]
    test_vals <- df$duree_p_emergence[df$statut == "test"]
    
    mean_ctrl <- mean(ctrl_vals, na.rm = TRUE)
    mean_test <- mean(test_vals, na.rm = TRUE)
    
      # Vérification normalité
    normal_ctrl <- shapiro.test(ctrl_vals)$p.value > 0.05
    normal_test <- shapiro.test(test_vals)$p.value > 0.05
    
    # Test statistique approprié
    if(normal_ctrl & normal_test){
        p <- t.test(duree_p_emergence ~ statut, data = df)$p.value
        test_used <- "t.test"
    } else {
        p <- wilcox.test(duree_p_emergence ~ statut, data = df, exact = FALSE)$p.value
        test_used <- "wilcox.test"
    }
    
    # Déterminer le label de signification
    if(p >= 0.05){
        label <- "ns"
    } else if(p < 0.001){
        label <- "***"
    } else if(p < 0.01){
        label <- "**"
    } else {
        label <- "*"
    }
    
     # Direction de l'effet
    Signif <- ifelse(p >= 0.05, "ns",
                     ifelse(mean(ctrl_vals) > mean(test_vals), "Ctrl > Inhib", "Inhib > Ctrl")
                    )
    
    tibble(durée = "Ponte_Emergence",
        mean_ctrl = round(mean_ctrl, 2),
        mean_test = round(mean_test, 2),
        p = round(p, 4),
        label = label,
        Signif = Signif,
        test_used = test_used
    )
}

# --- Calcul des p-values pour chaque inhibiteur ---
pval_global_duree_p_emergence <- data_for_analysis %>%
    dplyr::filter(nbre_adultes > 0) %>% 
    group_by(inhibitor) %>%
    summarise(
        get_pvalue_label(cur_data()),
        .groups = "drop"
    )
```
```{r}
Summary_pval_global_durations_cycle <- rbind(
  pval_global_duree_g_ponte,
  pval_global_duree_p_eclosion,
  pval_global_duree_eclo_nymphose,
  pval_global_duree_n_emergence,
  pval_global_duree_p_emergence
)
```

### MODEL DE REGRESSION #FS & KM 
#1.Charger les packages
```{r}
library(tidyverse)
library(lme4)
library(lmerTest)   # donne les p-values
library(emmeans)
library(ggplot2)
```

#2. Vérifier la structure de ton tableau
```{r}
data <- data %>%
  mutate(
    experience = as.factor(experience),
    inhibiteur = as.factor(inhibiteur),
    groupe     = as.factor(groupe)
  )
```

#3. MODÈLE MIXTE
```{r}
model <- lmer(valeur ~ inhibiteur * groupe + (1|experience), data = data)
summary(model)
```

#4. Comparaison entre inhibiteurs : EMMeans
```{r}
effets <- emmeans(model, ~ groupe | inhibiteur)
effets
pairs(effets)   # test vs control pour chaque inhibiteur
```

```{r}
emmeans(model, pairwise ~ inhibiteur | groupe)
```
```{r}
emmeans(model, pairwise ~ inhibiteur * groupe)
```

#5. EXTRACTION DIRECTE DE L’EFFET DE CHAQUE INHIBITEUR

```{r}
eff_traitement <- contrast(effets, method = "revpairwise") 
eff_traitement
```
#6. Graphique clair des effets

#Boxplot simple :
```{r}
ggplot(data, aes(x = inhibiteur, y = valeur, fill = groupe)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Effet des inhibiteurs sur le paramètre",
       y = "Valeur du paramètre",
       x = "Inhibiteur")
```

#Graphique des moyennes ajustées (avec EMMeans)
```{r}
df_emm <- as.data.frame(effets)

ggplot(df_emm, aes(x = inhibiteur, y = emmean, color = groupe)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = .1) +
  theme_minimal() +
  labs(title = "Effets ajustés (EMMeans)",
       y = "Moyenne ajustée",
       x = "Inhibiteur")
```

#7. Diagnostic du modèle
```{r}
par(mfrow=c(1,2))
plot(model)           # résidus vs ajustements
qqnorm(residuals(model)); qqline(residuals(model))
```



